Supplementary Information

### 2020-11-22 doc/manuscript/MEE_manuscript_SI.Rmd
NP feedback on appendix S3

To verify that our methods gave similar results to traditional methods, we conducted qualitative comparisons of our results to previous studies based on the same cores (Table `r qualitative_results_comparison`) and conducted a formal quantitative comparison for four species (Figs. `r all_traditional_comparisons`), as detailed below.

@@

Results from previous studies were compiled alongside results from the climate-only model if this study (Table `r qualitative_results_comparison`). 
Where previous studies examined numerous climate variables or time windows [e.g., @helcoski_growing_2019], we focus on those most relevant to our findings.

@@

We also conducted a formal comparison of our approach to conventional methods using identical tree-ring and climate data for four species: PSME (Cedar Breaks, Utah), ABAL ($\v{Z}$of$\'{i}$n), PIMA (Scotty Creek), and LITU (SCBI; Figs. `r all_traditional_comparisons`). 
These species were pre-selected for analysis because they have been well-studied in the past, and the species set was not manipulated after initiating the analysis.
For each species, we compared climate sensitivities for the top precipitation- and temperature- group variables, as identified in the main analysis.

@@

We note that despite designing the analyses to be as comparable as possible, one-to-one correspondence of *beta* coefficients is not necessarily expected for several reasons.
First, although the analysis time frame is standardized between the two approaches, the relative influence of each year will generally vary between the two approaches. 
The traditional approach, which all cores into a single residual chronology with one value per year, gives equal weighting to each year.
In contrast, the approach developed here effectively weights years relative to the number of cores available for each. 
Thus, where many younger trees are included in the analysis, the two approaches will effectively give different weights to the years included in the analysis period. 
In cases where climate-sensitivity differs between old and young trees, or where the climate and/or climate response changed substantially over the analysis time frame [e.g., at Scotty Creek; Fig. `r SC_traditional_comparison`; @sniderhan_growth_2016], this may lead to divergence of the climate sensitivities estimated by the two methods.

Second, traditional analysis methods (using ARSTAN) are designed to distill population-level variation to obtain the strongest possible climate signal for the reconstruction of past climate (**REF**), not to characterize climate responses on the individual level, where variation is inherently higher. 
While conversion of Pearson correlations to linear slopes *sensu* @charney_observed_2016 approximates climate responses, it does not provide an exact slope describing the relationship between individual-level or population mean growth and climate. 
This is because standardization of variance and averaging of individual-level residuals prior to the climate analysis fundamentally alters and obfuscates individual-level responses.

>>>>

To test whether our methods gave similar results to traditional methods, we conducted qualitative comparisons of our results to previous studies based on the same cores (Table `r qualitative_results_comparison`) and conducted a formal quantitative comparison for four species (Figs. `r all_traditional_comparisons`), as detailed below.

@@

Results from previous studies were compiled alongside results from the climate-only model in this study (Table `r qualitative_results_comparison`). 
Where previous studies examined numerous climate variables or time windows [e.g., @helcoski_growing_2019], we focus on those most relevant to our findings.

@@

We also conducted a formal comparison of our approach to conventional methods using identical tree-ring and climate data for four species: PSME (Cedar Breaks, Utah), ABAL ($\v{Z}$of$\'{i}$n), PIMA (Scotty Creek), and LITU (SCBI; Figs. `r all_traditional_comparisons`). 
These species were selected for analysis because they have been well-studied in the past.
For each species, we compared climate sensitivities for the top precipitation- and temperature- group variables, as identified in the main analysis.

@@

We note that despite designing the analyses to be as comparable as possible, one-to-one correspondence of *beta* coefficients is not necessarily expected for several reasons.
First, although the analysis time frame is standardized between the two approaches, the relative influence of each year will generally vary between the two approaches. 
The traditional approach, which all cores into a single residual chronology with one value per year, gives equal weighting to each year.
In contrast, under the approach developed here, the number records per year will vary across the analysis time frame, generally increasing over time as the younger trees enter the analysis. 
Thus, where many younger trees are included in the analysis, the two approaches will effectively give different weights to the years included in the analysis period. 
In cases where climate-sensitivity differs between old and young trees, or where the climate and/or climate response changed substantially over the analysis time frame [e.g., at Scotty Creek; Fig. `r SC_traditional_comparison`; @sniderhan_growth_2016], this may lead to divergence of the climate sensitivities estimated by the two methods.

Second, traditional analysis methods (using ARSTAN) were primarily designed to distill population-level variation to obtain the strongest possible climate signal for the reconstruction of past climate (**Cook & Kariustis **), not to characterize climate responses on the individual level, where variation is inherently higher. 
While conversion of Pearson correlations to linear slopes *sensu* @charney_observed_2016 approximates climate responses, it does not provide an exact slope describing the relationship between individual-level or population mean growth and climate. 
This is because standardization of variance and averaging of individual-level residuals prior to the climate analysis fundamentally alters and obfuscates individual-level responses.
